---
title: "DC area COVID hospitalizations"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(highcharter)

state_abb_list <- c("VA", "MD", "DC")
county_fips_list <- c(
  51510, # VA: Alexandria City
  51013, # VA: Arlington County
  51059, # VA: Fairfax County
  24003, # MD: Anne Arundel County 
  24017, # MD: Charles
  24031, # MD: Montgomery
  24033, # MD: Prince George's
  11001  # DC
)

county_pop <- read_csv("co-est2020.csv") %>% 
  mutate(county_fips = paste0(STATE, COUNTY)) %>% 
  select(county_fips, pop = POPESTIMATE2020, county_name = CTYNAME)

# source:
# https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u
county_patients <- read_csv("COVID-19_Reported_Patient_Impact_and_Hospital_Capacity_by_Facility.csv") %>% 
  rename(
    adults_sum = previous_day_admission_adult_covid_confirmed_7_day_sum,
    kids_sum = previous_day_admission_pediatric_covid_confirmed_7_day_sum,
    adults_reports = previous_day_admission_adult_covid_confirmed_7_day_coverage,
    kids_reports = previous_day_admission_pediatric_covid_confirmed_7_day_coverage
  ) %>% 
  # assume censored values and missings = 0
  mutate(across(adults_sum|kids_sum, ~ if_else(.x == -999999 | is.na(.x), 0, .x))) %>%
  mutate(adults_avg = if_else(
    adults_reports > 0,
    adults_sum / adults_reports,
    0
  )) %>% 
  mutate(kids_avg = if_else(
    kids_reports > 0,
    kids_sum / kids_reports,
    0
  )) %>% 
  group_by(county_fips = fips_code, week = collection_week) %>% 
  summarize(
    admits = sum(adults_avg + kids_avg),
    state_abb = first(state)
  ) %>% 
  ungroup() %>% 
  left_join(county_pop, by = "county_fips")

```

```{r, fig.width=10,fig.height=5}
county_patients %>% 
  filter(county_fips %in% county_fips_list) %>%
  group_by(week) %>% 
  summarize(across(admits|pop, sum)) %>% 
  ungroup() %>% 
  mutate(rate = admits / pop * 100000) %>% 
  mutate(rate = round(rate, 2)) %>% 
  hchart(
    "line", 
    hcaes(x = week, y = rate),
  ) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>Eight-county area</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(
    title = FALSE,
    tickInterval = 0.5,
    labels = list(format = "{value:.1f}")
  ) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "Eight-county area confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("area_level")

county_patients %>% 
  filter(county_fips %in% county_fips_list) %>% 
  mutate(rate = admits / pop * 100000) %>% 
  mutate(rate = round(rate, 2)) %>% 
  hchart(
    "line", 
    hcaes(x = week, y = rate, group = county_name),
    showInLegend = TRUE
  ) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>{series.name}</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(title = FALSE) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "County-level confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("county_level")

us_pop <- read_csv("co-est2020.csv") %>% 
  filter(SUMLEV == "040") %>% 
  summarize(sum(POPESTIMATE2020)) %>% 
  pull()

us_overall <- county_patients %>% 
  group_by(week) %>% 
  summarize(admits = sum(admits)) %>% 
  mutate(pop = us_pop) %>% 
  ungroup() %>% 
  mutate(state_abb = "US overall")

county_patients %>% 
  filter(state_abb %in% state_abb_list) %>% 
  group_by(state_abb, week) %>% 
  summarize(across(admits|pop, sum)) %>% 
  ungroup() %>% 
  bind_rows(us_overall) %>% 
  mutate(state_abb = factor(state_abb, levels = c("DC", "MD", "VA", "US overall"))) %>% 
  mutate(rate = admits / pop * 100000) %>% 
  mutate(rate = round(rate, 2)) %>% 
  hchart(
    "line", 
    hcaes(x = week, y = rate, group = state_abb),
    showInLegend = TRUE
  ) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>{series.name}</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(title = FALSE) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "State-level confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("state_level")
```


The figures show daily averages, by week, of the number of adult and pediatric patients admitted to inpatient beds on the previous day who had confirmed COVID-19 at the time of admission. These averages are expressed as rates per 100,000 residents. The date of week above refers to start of the week: "February 4" means the week beginning February 4 and ending February 10, inclusive.

The hospital admission data is facility-level data from healthdata.gov [here](https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u). To calculate the rate for a given area, first calculate the average daily adult and pediatric admissions in each facility in that area: divide the facility-reported weekly sum of patients by the facility-reported weekly number of reports. Then divide the average admissions by the Census Vintage 2020 estimated resident population available [here](https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-counties-total.html). The code that does the analysis and creates this page is at https://github.com/Economic/dc_hospitalizations.

The results are broadly similar, but not exactly the same, as CDC statistics presented in a less accessible form [here](https://covid.cdc.gov/covid-data-tracker/#new-hospital-admissions). One coverage difference is the CDC counts might include Veterans Affairs and some other facilities excluded in the healthdata.gov data.

Last updated February 24, 2022 using vintage February 21, 2022 11:35 AM EST hospitalization data.