---
title: "DC area COVID hospitalizations"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(highcharter)
library(timetk)
library(kableExtra)

state_abb_list <- c("VA", "MD", "DC")
county_data <- tribble(
  ~county_name, ~county_fips, ~lodes_emp,
  "Alexandria City", 51510, 20625,
  "Anne Arundel County", 24003, 13576,
  "Arlington County", 51013, 36506,
  "Charles County", 24017, 14663,
  "District of Columbia", 11001, 211689,
  "Fairfax County", 51059, 73105,
  "Montgomery County", 24031, 90013,
  "Prince George's County", 24033, 129304
) %>% 
  mutate(
    county_fips = as.character(county_fips),
    emp_share = lodes_emp / sum(lodes_emp)
  )

county_fips_list <- county_data %>% 
  pull(county_fips)

county_pop <- read_csv("co-est2020.csv") %>% 
  mutate(county_fips = paste0(STATE, COUNTY)) %>% 
  select(county_fips, pop = POPESTIMATE2020, county_name = CTYNAME)

# source:
# https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u
county_patients <- read_csv("COVID-19_Reported_Patient_Impact_and_Hospital_Capacity_by_Facility.csv") %>% 
  rename(
    adults_sum = previous_day_admission_adult_covid_confirmed_7_day_sum,
    kids_sum = previous_day_admission_pediatric_covid_confirmed_7_day_sum,
    adults_reports = previous_day_admission_adult_covid_confirmed_7_day_coverage,
    kids_reports = previous_day_admission_pediatric_covid_confirmed_7_day_coverage
  ) %>% 
  # assume censored values and missings = 0
  mutate(across(adults_sum|kids_sum, ~ if_else(.x == -999999 | is.na(.x), 0, .x))) %>%
  mutate(adults_avg = if_else(
    adults_reports > 0,
    adults_sum / adults_reports,
    0
  )) %>% 
  mutate(kids_avg = if_else(
    kids_reports > 0,
    kids_sum / kids_reports,
    0
  )) %>% 
  group_by(county_fips = fips_code, week = collection_week) %>% 
  summarize(
    admits = sum(adults_avg + kids_avg),
    state_abb = first(state)
  ) %>% 
  ungroup() %>% 
  left_join(county_pop, by = "county_fips")

# county cases
# source:
# https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-Community-Levels-by-County/3nnm-4jni
county_cases <- read_csv("United_States_COVID-19_Community_Levels_by_County.csv") %>%
  filter(county_fips %in% county_fips_list) %>%
  select(date_updated, county, county_fips, county_population, state, covid_cases_per_100k) %>% 
  mutate(pop_share = county_population / sum(county_population), .by = date_updated) %>%
  left_join(county_data) %>% 
  arrange(date_updated)

```

```{r, fig.width=10,fig.height=6}
county_patients %>% 
  filter(county_fips %in% county_fips_list) %>%
  full_join(county_data, by = "county_fips") %>% 
  summarize(rate_emp_weight = sum(admits / pop * emp_share),.by = week) %>% 
  mutate(rate_emp_weight = rate_emp_weight * 100000) %>% 
  mutate(rate_emp_weight = round(rate_emp_weight, 2)) %>% 
  tk_xts(rate_emp_weight, date_var = week) %>% 
  hchart() %>% 
  hc_rangeSelector(enabled = FALSE) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>Eight-county area</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(
    title = FALSE,
    tickInterval = 0.5,
    labels = list(format = "{value:.1f}"),
    opposite = FALSE,
    plotLines = list(
      list(
        label = list(text = "1.75"),
        color = "#f55f0f",
        width = 1,
        value = 1.75,
        # the zIndex is used to put the label text over the grid lines 
        zIndex = 1
        )
    )  
  ) %>% 
  hc_plotOptions(
    series = list(
      marker = list(enabled = TRUE, radius = 4),
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "Eight-county area confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents, weighted by county employment shares",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("area_level") %>% 
  hc_exporting(
      enabled = TRUE, # always enabled
      filename = "eight-county-hospitalizations"
      )
```

```{r, fig.width=10,fig.height=6}
county_cases %>% 
  summarize(
    wtd_cases_per_100k = sum(covid_cases_per_100k * emp_share), 
    .by = date_updated) %>% 
  pivot_longer(-date_updated) %>% 
    mutate(name = case_when(
    name == "wtd_cases_per_100k" ~ "Average weekly cases"
  )) %>% 
  hchart(
    "line", 
    hcaes(x = date_updated, y = value, group = name),
    showInLegend = TRUE
  ) %>%
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>{series.name}</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(
    title = FALSE,
    tickInterval = 0.5,
    labels = list(format = "{value:.1f}"),
    plotLines = list(
      list(
        label = list(text = "200"),
        color = "#f55f0f",
        width = 1,
        value = 200,
        # the zIndex is used to put the label text over the grid lines 
        zIndex = 1
      )
    )
  ) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "Eight-county area average COVID cases, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Weekly cases per 100,000 residents, weighted by county employment shares",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("emp-weighted-series") %>% 
  hc_exporting(
    enabled = TRUE, # always enabled
    filename = "emp-weighted-cases"
  )
  
```

The data displayed in the two charts above is weighted by employment shares, where the employment share of an county is that county's share of total employment in DC coming from the eight counties, as reported by the Census OnTheMap [tool](https://onthemap.ces.census.gov/) for 2019.

Each county's total employment in DC and associated share is the following:

```{r}
county_data %>%
  column_to_rownames("county_name") %>% 
  mutate(
    emp_share = scales::label_percent(accuracy = 0.1)(emp_share),
    lodes_emp = scales::label_comma()(lodes_emp)
  ) %>% 
  select("2019 employment in DC" = lodes_emp, "Share" = emp_share) %>% 
  kbl(align = "r") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

```{r, fig.width=10,fig.height=5}
county_patients %>% 
  filter(county_fips %in% county_fips_list) %>% 
  mutate(rate = admits / pop * 100000) %>% 
  mutate(rate = round(rate, 2)) %>% 
  hchart(
    "line", 
    hcaes(x = week, y = rate, group = county_name),
    showInLegend = TRUE
  ) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>{series.name}</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(title = FALSE) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "County-level confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("county_level") %>% 
  hc_exporting(
    enabled = TRUE, # always enabled
    filename = "county-level-hospitalizations"
  )
```

```{r, fig.width=10,fig.height=5}
us_pop <- read_csv("co-est2020.csv") %>% 
  filter(SUMLEV == "040") %>% 
  summarize(sum(POPESTIMATE2020)) %>% 
  pull()

us_overall <- county_patients %>% 
  group_by(week) %>% 
  summarize(admits = sum(admits)) %>% 
  mutate(pop = us_pop) %>% 
  ungroup() %>% 
  mutate(state_abb = "US overall")

county_patients %>% 
  filter(state_abb %in% state_abb_list) %>% 
  group_by(state_abb, week) %>% 
  summarize(across(admits|pop, sum)) %>% 
  ungroup() %>% 
  bind_rows(us_overall) %>% 
  mutate(state_abb = factor(state_abb, levels = c("DC", "MD", "VA", "US overall"))) %>% 
  mutate(rate = admits / pop * 100000) %>% 
  mutate(rate = round(rate, 2)) %>% 
  hchart(
    "line", 
    hcaes(x = week, y = rate, group = state_abb),
    showInLegend = TRUE
  ) %>% 
  hc_tooltip(
    crosshairs = TRUE, 
    headerFormat = "<b>{series.name}</b><br>",
    pointFormat = "{point.x:%B %d, %Y}<br>{point.y:.2f}"
  ) %>% 
  hc_xAxis(title = FALSE) %>% 
  hc_yAxis(title = FALSE) %>% 
  hc_plotOptions(
    series = list(
      states = list(inactive = list(opacity = 1))
    )
  ) %>% 
  hc_title(
    text = "State-level confirmed COVID hospital admissions, by week",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "26px")
  ) %>% 
  hc_subtitle(
    text = "Daily average per 100,000 residents",
    align = "left",
    style = list(fontWeight = "bold", fontSize = "20px")
  ) %>% 
  hc_elementId("state_level") %>% 
  hc_exporting(
    enabled = TRUE, # always enabled
    filename = "state-level-hospitalizations"
  )
```

The figures show daily averages, by week, of the number of adult and pediatric patients admitted to inpatient beds on the previous day who had confirmed COVID-19 at the time of admission. These averages are expressed as rates per 100,000 residents. The date of week above refers to start of the week; for example, "February 4" is the week beginning February 4 and ending February 10, inclusive.

The hospital admission data is facility-level data from healthdata.gov [here](https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u). To calculate the rate for a given area, first calculate the average daily adult and pediatric admissions in each facility in that area: divide the facility-reported weekly sum of patients by the facility-reported weekly number of reports. Then divide the average admissions by the Census Vintage 2020 estimated resident population available [here](https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-counties-total.html). The code that does the analysis and creates this page is at https://github.com/Economic/dc_hospitalizations.

The results are broadly similar, but not exactly the same, as CDC statistics presented in a less accessible form [here](https://covid.cdc.gov/covid-data-tracker/#new-hospital-admissions). One coverage difference is the CDC counts might include Veterans Affairs and some other facilities excluded in the healthdata.gov data.

The figures above relating to number of cases show weekly (7-day total) covid cases per 100,000 residents in the DC area.The data for [weekly cases](https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-Community-Levels-by-County/3nnm-4jni) are available at the county level and are updated weekly on thursdays.


Last updated June 27, 2023 using vintage June 26, 2023 hospitalization data and June 1, 2023 case data. NOTE: Reporting of Aggregate Case and Death Count data was discontinued May 11, 2023, with the expiration of the COVID-19 public health emergency declaration. 

